---
title: "Structuration of the vertebrate diversity in metropolitan France"
author: "RL"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE,comment = FALSE,prompt = FALSE,error = FALSE, message = FALSE)
```




```{r}
library(data.table)
library(ggplot2)
library(magrittr)
library(kableExtra)
source("../functions/children_of_CD_NOM_fb.R")
```

# TAXREF

We load the last version of the taxref database, TAXREFv16 2022. 
We only keep the FR column for metropolitan France and the lines corresponding to vertebrate species (without the synonyms).

```{r}
taxref <- fread('../data/TAXREF_v16_2022/TAXREFv16.txt',
  encoding="UTF-8",select = c("CD_NOM", "CD_REF", "CD_SUP", "REGNE", "PHYLUM",
  "CLASSE", "ORDRE", "FAMILLE", "GROUP1_INPN", "GROUP2_INPN", "RANG", "HABITAT",
  "FR"))
v_class_verterbrate <- c("Actinopterygii","Amphibia","Aves","Dipneusti","Elasmobranchii","Holocephali","Mammalia","Myxini","Petromyzonti")
v_ordre_vertebrate <- c("Chelonii", "Squamata", "Crocodylia")
taxref <- taxref[CD_NOM == CD_REF & RANG == "ES" & (CLASSE %in%v_class_verterbrate | ORDRE %in% v_ordre_vertebrate),]
#kable(head(taxref),booktabs = TRUE)
```

We keep only some status 

```{r}
status <- fread("../data/TAXREF_v16_2022/statuts_note.csv",encoding="Latin-1")
v_keep_status <- c("P", "E", "S", "C", "I", "J")
v_natif <- c("P","E","S")
v_introduit <- c("I","J")
status[,`:=` (KEEP = STATUT %in% v_keep_status,
              NATIF_INTRODUIT = ifelse(STATUT %in% v_natif, "NATIVE", ifelse(STATUT %in% v_introduit, "INTRODUCED",NA)))]
setorder(x=status,-KEEP,ORDRE)
setnames(status, c("STATUT","DESCRIPTION"),c("STATUS","DESCRIPTION_STATUS"))

status <- status[,.(KEEP,STATUS,NATIF_INTRODUIT,DESCRIPTION_STATUS)]
k_status <- kable(status,format= "html",full_width = FALSE)
kable_styling(k_status, "striped", position = "left", font_size = 12)
```




```{r}
taxref <- merge(taxref,status[KEEP==TRUE,],by.x = "FR",by.y="STATUS")
setnames(taxref,"FR","STATUS_FR")
```


The data set : 





```{r}
habitat <- fread("../data/TAXREF_v16_2022/habitats_note.csv")
habitat <- habitat[,.(HABITAT,HABITAT_name)]
taxref <- merge(taxref,habitat,by="HABITAT")
```

```{r}
the_col <- c("CD_NOM", "CLASSE", "ORDRE", "FAMILLE", "GROUP1_INPN", "GROUP2_INPN","HABITAT","HABITAT_name",
  "STATUS_FR","NATIF_INTRODUIT","DESCRIPTION_STATUS")
taxref <- taxref[,..the_col]
#head(taxref) %>% 
#  kable_paper("striped") #%>%
#  scroll_box(width = "100%", height = "200px")

```

# Number of species 
## All species

```{r}

print(nrow(taxref))

print(taxref[,.(nb = .N),by= GROUP2_INPN])


```
## Continental habitats

## Number of species

```{r}
vec_cont <- c(2:5,7:8)

vec_cont <- c(2:5,7:8)
taxref_cont <- taxref[HABITAT %in% vec_cont,]
taxref_cont[HABITAT == 5, HABITAT_name := "Land"]
taxref_cont[HABITAT == 4, HABITAT_name := "Freshwater"]
taxref_cont[HABITAT == 7, HABITAT_name := "Continental (or)"]
taxref_cont[HABITAT == 8, HABITAT_name := "Continental (and)"]

print(nrow(taxref_cont))

print(taxref_cont[,.(nb=.N),by= .(GROUP2_INPN)])

taxref_cont <- taxref[HABITAT %in% vec_cont,]
gg_tax_ordre <- taxref_cont[,.(nb = .N),by = .(GROUP2_INPN,ORDRE)]
setorder(gg_tax_ordre,-nb)
gg_tax_ordre[,ORDRE := factor(ORDRE, level = ORDRE)]
gg <- ggplot(gg_tax_ordre,aes(x=ORDRE,y=nb,fill=GROUP2_INPN))+ geom_bar(stat="identity")
gg <- gg + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
print(gg)

```

### habitat
```{r}


gg_tax <- taxref_cont[,.(nb = .N),by= .(GROUP2_INPN,HABITAT_name)]
gg <- ggplot(data=gg_tax,aes(y=nb,x=GROUP2_INPN,fill=GROUP2_INPN)) + facet_grid(~HABITAT_name,scales="free")
gg <- gg + geom_bar(stat="identity", width=1) 
print(gg)
```

### introduced species


```{r}

gg_tax <- taxref_cont[!is.na(NATIF_INTRODUIT),.(nb = .N),by= .(GROUP2_INPN,NATIF_INTRODUIT)]
gg <- ggplot(data=gg_tax,aes(y=nb,x=GROUP2_INPN,fill=GROUP2_INPN)) + facet_wrap(.~NATIF_INTRODUIT,scales="free")
gg <- gg + geom_bar(stat="identity", width=1) 
print(gg)
```



```{r}
gg_tax <- taxref_cont[!is.na(NATIF_INTRODUIT),.(nb = .N),by= .(GROUP2_INPN,HABITAT_name,NATIF_INTRODUIT)]
gg <- ggplot(data=gg_tax,aes(y=nb,x=GROUP2_INPN,fill=GROUP2_INPN)) + facet_grid(NATIF_INTRODUIT~HABITAT_name,scales="free")
gg <- gg + geom_bar(stat="identity", width=1) 
print(gg)
```




## Marine habitats

## Number of species

```{r}
vec_marine <- c(1,4,5)


taxref_marine <- taxref[HABITAT %in% vec_marine,]
taxref_marine[HABITAT == 5, HABITAT_name := "Marine"]
taxref_marine[HABITAT == 4, HABITAT_name := "Marine"]
taxref_marine[HABITAT == 7, HABITAT_name := "marineinental (or)"]
taxref_marine[HABITAT == 8, HABITAT_name := "marineinental (and)"]

print(nrow(taxref_marine))

print(taxref_marine[,.(nb=.N),by= .(GROUP2_INPN)])

taxref_marine <- taxref[HABITAT %in% vec_marine,]
gg_tax_ordre <- taxref_marine[,.(nb = .N),by = .(GROUP2_INPN,ORDRE)]
setorder(gg_tax_ordre,-nb)
gg_tax_ordre[,ORDRE := factor(ORDRE, level = ORDRE)]
gg <- ggplot(gg_tax_ordre,aes(x=ORDRE,y=nb,fill=GROUP2_INPN))+ geom_bar(stat="identity")
gg <- gg + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
print(gg)

```

### habitat
```{r}


gg_tax <- taxref_marine[,.(nb = .N),by= .(GROUP2_INPN,HABITAT_name)]
gg <- ggplot(data=gg_tax,aes(y=nb,x=GROUP2_INPN,fill=GROUP2_INPN)) + facet_grid(~HABITAT_name,scales="free")
gg <- gg + geom_bar(stat="identity", width=1) 
print(gg)
```

### introduced species


```{r}

gg_tax <- taxref_marine[!is.na(NATIF_INTRODUIT),.(nb = .N),by= .(GROUP2_INPN,NATIF_INTRODUIT)]
gg <- ggplot(data=gg_tax,aes(y=nb,x=GROUP2_INPN,fill=GROUP2_INPN)) + facet_wrap(.~NATIF_INTRODUIT,scales="free")
gg <- gg + geom_bar(stat="identity", width=1) 
print(gg)
```



```{r}
gg_tax <- taxref_marine[!is.na(NATIF_INTRODUIT),.(nb = .N),by= .(GROUP2_INPN,HABITAT_name,NATIF_INTRODUIT)]
gg <- ggplot(data=gg_tax,aes(y=nb,x=GROUP2_INPN,fill=GROUP2_INPN)) + facet_grid(NATIF_INTRODUIT~HABITAT_name,scales="free")
gg <- gg + geom_bar(stat="identity", width=1) 
print(gg)
```

# Next step 
 - spatial distribution (number of species in bio-climatic region)
 - functional traits (taxref)
    - diet 
    - mass
    - migration strategy
 - conservation status (IUCN)
 
 

